---
title: ICA Experiment
date: 2022-12-18
---

The explanation of ICA is [Independent Component Analysis]()

## Central limit theorem


The distribution of sum of independent random variables tends toward a Gaussian distribution, under certain condition.

It specific certain condition, we here restrict to Independent and identically distributed(i.i.d.) condition. 

So the experience here we will sum up two iid random variable and show there distribution.

### Uniform distribution

```{r}
# set.seed(1)
n_sample = 10000
n_vector = 3
x1 = runif(n_sample)
rvs = runif(n_sample * n_vector)
dim(rvs) = c(n_sample, n_vector)
sum_rvs = apply(rvs, 1, mean)
mean = 0.5
sd = sqrt(1/12)/sqrt(n_vector)
p0 = hist(x1, xlab = 'x', freq=FALSE, density = 20, nclass = 40, prob = TRUE)
p1 = hist(sum_rvs, xlab = 'x', freq=FALSE, density = 20, nclass = 40, prob = TRUE)
gaussian = rnorm(n_sample, mean=mean, sd= sd)
p2 = hist(gaussian, xlab = 'x', freq=FALSE, density = 20, nclass = 80, prob = TRUE)
plot(p1, col=rgb(0,0,1,1/5), xlim=c(0,1), freq = FALSE)
plot(p2, col=rgb(1,0,0,1/5), xlim=c(0,1), freq = FALSE, add = T)  # second
# rvs = append(rvs, runif(n_sample))


```
### Exponential Distribution

```{r}
n_sample = 10000
n_vector = 20
lambda = 1
x1 = rexp(n_sample, lambda)
rvs = rexp(n_sample * n_vector, lambda)
dim(rvs) = c(n_sample, n_vector)
sum_rvs = apply(rvs, 1, mean)
mean = 1/lambda
sd = 1/lambda/sqrt(n_vector)
mean = mean
p0 = hist(x1, xlab = 'x', freq=FALSE, density = 20, nclass = 80, prob = TRUE)
p1 = hist(sum_rvs, xlab = 'x', density = 20, nclass = 80, prob = TRUE)
gaussian = rnorm(n_sample, mean=mean, sd= sd)
p2 = hist(gaussian, xlab = 'x', density = 20, nclass = 80, prob = TRUE)
plot(p1, col=rgb(0,0,1,1/5), xlim=c(-2,2))
plot(p2, col=rgb(1,0,0,1/5), xlim=c(-2,2), add = T)  # second


```



### Non Gaussian is indepence

Previous experient shows the central limit theorem. The distribution of sum of independent random variables tends toward a Gaussian distribution, under certain condition.

So the ICA is maximize the nongaussian of the input signal. 

#### Measure of nongaussian

We using following equation to measure the 

$$
\begin{align} 
J(y) \propto[E\{G(y)\}-E\{G(v)\}]^2 \\
G_1(u)=\frac{1}{a_1} \log \cosh a_1 u, \quad G_2(u)=-\exp \left(-u^2 / 2\right) \\
g_1(u)=\tanh \left(a_1 u\right) \\ 
g_2(u)=u \exp \left(-u^2 / 2\right)
\end{align}
$$

define $g_1$, $g_2$, $g_1'$ and $g_2'$
```{r}
a1 = 1;
g1 <- function(u) {
  return (tanh(a1 * u))
}
g1_prime <- function(u) {
  return (a1 * 1-tanh(a1 * u)^2)
}
g2 <- function(u){
  return (u * exp(-u*u/2))
}
g2_prime <- function(u) {
  return (exp(u^2/2) + u^2*exp(u^2/2))
}

```

Generate observe signal x(t) from siganl s(t)
Here we define $s_1(t) = sin(t), s_2(t) = t - [t]$ where [t] is the maximun integer less than $t$.

```{r}
t = seq(0, 50, 0.001)

# generate s1
s1 = sin(t)
plot(t, s1)
```

Generate s2
```{r}
s2_func <- function(t1) {
  t1= t1 / 10
  tmp = t1 - floor(t1)
  if (tmp <= 0.8) {
    return(tmp)
  }
  else {
    return (0.8 * (1-tmp)/0.2)
  }
}
# s2 = t - floor(t)
# s2= s2_func(t)
s2 = unlist((lapply(t, s2_func)))
plot(t, s2)
```
Generate $x_1$ and $x_2$
```{r}
A = c(1, 0.1, 0.01, 1)
dim(A) = c(2,2)
x1 = A[1,1] * s1 + A[1,2] * s2
x2 = A[2,1] * s1 + A[2,2] * s2
X = c(x1,x2)
dim(X) = c(length(t),2)
X = t(X)

# plot x1 and x2
plot(t, X[1,])
plot(t, X[2,])
```


Preprocess
```{r}
# Remove mean
n = length(t)
X = X - apply(X, 1, mean)

# Whiten
C = 1/n * (X %*% t(X))
r = eigen(C)
E = r$vectors
D = diag(r$values)
rsqrt_D = diag(1/r$values)
x_tilda = E %*% rsqrt_D %*%t(E) %*% X

x = x_tilda

```

Run one unit ICA
```{r}

w = rnorm(2)
dim(w) = c(2,1)

for (i in 1:50) {
  w_tmp = 1/n * (x %*% t(g1(t(w) %*% x))) - mean(g1_prime(t(w) %*% x)) * w
  w = w_tmp / norm(w_tmp, type="2")
}

s1_bar = t(w) %*% x
plot(t, s1_bar)


```














